---
title: "Data Basics (solutions)"
output:
  github_document: 
    df_print: tibble
  html_document:
    df_print: paged
---

<!-- This file by Charlotte Wickham is licensed under a Creative Commons Attribution 4.0 International License. -->

```{r setup}
library(tidyverse)
library(haven)
library(readxl)
```

# Tabular Data

You might not see much difference between these two in this notebook, but one is a tibble and one is a data frame:
```{r}
mpg
```

```{r}
quakes
```

## Your turn 1

Take a look at these two datasets, by typing their names into the console:
* quakes
* mpg

Printing a data frame in R will print all columns and all the rows (until meeting some maximum printing value).  In contrast, printing a tibble will print only as many columns as will fit on your screen and only 10 rows of the the data.  Tibbles are a special type of data frame and have a few other special behaviours too.  In this example, mpg is a tibble and quakes is a data frame.

## Your turn 2

Try running each line one by one, with Cmd/Ctrl + r.  What do they do?

```{r}
dim(x = mpg)     
names(x = mpg)   
glimpse(x = mpg) 
View(x = mpg)  
```

* `dim()` prints the dimensions, number of observations then number of variables
* `names()` prints the column names
* `glimpse()` gives an overview of the data, including all variable names, their types, and the first few entries.
* `View()` opens up a GUI viewer.

## Your turn 3

How many rows of data are in quakes?

You could use `dim()`, `glimpse()` or `View()`:
```{r}
dim(quakes)

# or more specifically
nrow(quakes)
```

What are the names of the variables in quakes?

```{r}
names(quakes)
```

## Your turn 4

Run the chunk to get help on `mpg`. What is this data?
```{r}
?mpg
```

> "This dataset contains a subset of the fuel economy data that the EPA makes available on http://fueleconomy.gov. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car."

# Vector Data

## Your turn 5

Take another look at mpg.  What kind of data is in each column?

```{r}
mpg # and view in the notebook
glimpse(mpg)
```

There are character, double and integer variables.

# Importing Data

## Your Turn 6

Take a look in the data/ directory in the project folder.

Try reading in each of:

* `deaths.xls`
* `nimbus.csv`
* `iris.sav`


```{r}
deaths <- read_excel("data/deaths.xls")
deaths
```

```{r}
iris <- read_spss("data/iris.sav")
iris
```

```{r}
nimbus <- read_csv("data/nimbus.csv")
nimbus
```

## Your Turn 7

Can you see what is wrong with the Excel file when it is imported?

Scan the Arguments section of ?read_excel, can you find an argument that might help? Try it!

```{r}
deaths
```

Looks like the first few rows don't contain data, but some multi-column notations.  There's also some at the end.

One solution is to use the arguments `skip` and `n_max`:
```{r}
deaths <- read_excel("data/deaths.xls", skip = 4, n_max = 10)
deaths
```
 
 Another is to use the `range` argument:
```{r}
deaths <- read_excel("data/deaths.xls", range = "A5:F15")
deaths

```
 
