---
title: "Tidying New House Price Index"
output: html_notebook
---

```{r setup}
library(tidyverse)
```

# Importing raw(er) data

```{r}
housing_raw <- read_csv("data/housing_raw.csv")
```

# Tidying data

## Your turn

Brainstorm with neighbor: What needs to happen to `housing_raw` to get it in a form to plot?

```{r}
housing_raw
```

## Your turn

Try to tidy `housing_raw` don't forget your cheatsheet.

```{r}
housing_reshaped <- housing_raw %>% 
  gather(key = "time", value = "hpi", 
    -Geography, -`New housing price indexes`) %>%
  spread(`New housing price indexes`, hpi) 

housing_reshaped
```

Data looks to be in the right shape, but plotting shows some problems:

```{r}
ggplot(housing_reshaped) +
  geom_line(aes(x = time, y = `Total (house and land)`, color = Geography))
```

Problems:

* `time` is a character string, rather than a date
* `Geography` still contatins too many regions
* and it's painful to refer to a column with spaces in it.

Fixes:
```{r}
housing_bc <- housing_reshaped %>%
  mutate(date = lubridate::parse_date_time(time, order = "my")) %>%
  filter(Geography %in% c("Canada", "British Columbia (12)")) %>%
  rename(total = "Total (house and land)",
         land = "Land only",
         house = "House only",
         geography = Geography)
```


Then we can plot:
```{r}
ggplot(housing_bc) +
  geom_line(aes(x = date, y = total, color = geography))
```

